In this thesis, I am building a market simulator that is based on a LOB queuing model to test various optimal trade execution strategies. This process involves developing a model that accurately reflects real world dynamics of the LOB, estimating the parameters, and writing a program for simulation that can take decisions by an agent and update in real time. I am using data from trades on Ethereum Classic cryptocurrency from the Coinbase Pro exchange. My progress thus far can be summarized as follows:

\begin{itemize}
\item \textbf{Software}: I have organized my code in a github repository
\item \textbf{Refined Data Collection}: I have written code to collect and store updates. I have collected several days worth of data in the months of December and January.
\item \textbf{Data Processing}: I have written code to maintain an internal LOB for analysis
\item \textbf{Data Analysis and Model Development}: I started with a queuing model developed by \cite{A6} and modified it based on what I have analyzed in the data. This includes:
    \begin{itemize}
    \item Verifying that inter-arrival times are approximately exponential and estimating the average rate of arrival
    \item Estimating average event sizes and allowing for variable size events
    \item Finding correlations between the Poisson arrivals
    \end{itemize}
\item \textbf{Simulation Algorithm}: I have developed the simulation algorithm and begun writing preliminary code for parts of it (such as generating correlated Poisson arrivals using copulas and backward simulation)
\item \textbf{Thesis Writing}: I have written rough drafts of the introduction, literature review, data description, and experimental procedure for the data analysis described above. I also have the boilerplate for the final product written in LaTeX.
\end{itemize}